---
title: "QMM Assignment 2 Module 4"
author: "Pasala Dileep Kumar_811350727"
date: "2024-09-22"
output: html_document
---

Here as per the problem, profit should be maximized by utilizing excess capacity and by producing the quantities of new product with combination of product sizes.

**Assumptions**:
Let,

The quantity of product to be produced be “V”

Sizes of product to be produced be – “a” for large, “b” for medium, “c” for small.

Profit to company be “P”

❖ **Decision variables**:

Quantity of large size product produced per day in plant 1: “Va1”

Quantity of medium size product produced per day in plant 1: “Vb1”

Quantity of small size product produced per day in plant 1: “Vc1”

Quantity of large size product produced per day in plant 2: “Va2”

Quantity of medium size product produced per day in plant 2: “Vb2”

Quantity of small size product produced per day in plant 2: “Vc2”

Quantity of large size product produced per day in plant 3: “Va3”

Quantity of medium size product produced per day in plant 3: “Vb3”

Quantity of small size product produced per day in plant 3: “Vc3”


❖ **Objective function**:

**Maximise P = 420( Va1 + Va2 + Va3) + 360( Vb1 + Vb2 + Vb3) + 300( Vc1 + Vc2 + Vc3)**

❖ **Constraints**: 
5 types of constraints 

**i. Excess capacity constraint**:

➢ As per plant 1: Va1 + Vb1 + Vc1 ≤ 750

➢ As per plant 2: Va2 + Vb2 + Vc2 ≤ 900

➢ As per plant 3: Va3 + Vb3 + Vc3≤ 450

**ii. Space constraint for in process storage**:

➢ with respect to plant 1: 20Va1 + 15Vb1 + 12Vc1 ≤ 13000

➢ with respect to plant 2: 20Va2 + 15Vb2 + 12Vc2 ≤ 12000

➢ with respect to plant 3: 20Va3 + 15Vb3 + 12Vc3 ≤ 5000

**iii. Forecast of sales constraint**:

➢ for large size products(a): Va1 + Va2 + Va3 ≤ 900

➢ for medium size products(b): Vb1 + Vb2 + Vb3 ≤ 1200

➢ for small size products(c): Vc1 + Vc2 + Vc3 ≤ 750

**iv. Utilization of equal percentage of excess capacity constraint**:

➢ plant1: (Va1 + Vb1 + Vc1)/750 = e

➢ plant2: (Va2 + Vb2 + Vc2)/900= e

➢ plant3: (Va3 + Vb3 + Vc3)/450 = e

➢ moreover, 0 ≤ e ≤ 1

**v. Non negativity constraint**:

decision variable to be ≥ 0

i.e. Va1 ≥ 0, Vb1 ≥ 0, Vc1≥ 0, Va2 ≥ 0, Vb2 ≥ 0, Vc2 ≥ 0, Va3 ≥ 0, Vb3 ≥ 0, Vc3 ≥ 0

**Mathematical formulation**:

Maximise P = 420( Va1 + Va2 + Va3) + 360( Vb1 + Vb2 + Vb3) + 300( Vc1 + Vc2 + Vc3)

given that,

Va1 + Vb1 + Vc1 ≤ 750

Va2 + Vb2 + Vc2 ≤ 900

Va3 + Vb3 + Vc3≤ 450

20Va1 + 15Vb1 + 12Vc1 ≤ 13000

20Va2 + 15Vb2 + 12Vc2 ≤ 12000

20Va3 + 15Vb3 + 12Vc3 ≤ 5000

Va1 + Va2 + Va3 ≤ 900

Vb1 + Vb2 + Vb3 ≤ 1200

Vc1 + Vc2 + Vc3 ≤ 750

(Va1 + Vb1 + Vc1)/750 = e

(Va2 + Vb2 + Vc2)/900= e

(Va3 + Vb3 + Vc3)/450 = e

0 ≤ e ≤ 1

Va1 ≥ 0, Vb1 ≥ 0, Vc1≥ 0, Va2 ≥ 0, Vb2 ≥ 0, Vc2 ≥ 0, Va3 ≥ 0, Vb3 ≥ 0, Vc3 ≥ 0

Rewriting the above constraints as below:

Va1 + Vb1 + Vc1 + 0Va2 + 0Vb2 + 0Vc2 + 0Va3 + 0Vb3 + 0Vc3 ≤ 750

0Va1 + 0Vb1 + 0Vc1 + Va2 + Vb2 + Vc2 + 0Va3 + 0Vb3 + 0Vc3 ≤ 900

0Va1 + 0Vb1 + 0Vc1 + 0Va2 + 0Vb2 + 0Vc2 + Va3 + Vb3 + Vc3 ≤ 450

20Va1 + 15Vb1 + 12Vc1 + 0Va2 + 0Vb2 + 0Vc2 + 0Va3 + 0Vb3 + 0Vc3 ≤ 13000

0Va1 + 0Vb1 + 0Vc1 + 20Va2 + 15Vb2 + 12Vc2 + 0Va3 + 0Vb3 + 0Vc3 ≤ 12000

0Va1 + 0Vb1 + 0Vc1 + 0Va2 + 0Vb2 + 0Vc2 + 20Va3 + 15Vb3 + 12Vc3 ≤ 5000

Va1 + 0Vb1 + 0Vc1 + Va2 + 0Vb2 + 0Vc2 + Va3 + 0Vb3 + 0Vc3 ≤ 900

0Va1 + Vb1 + 0Vc1 + 0Va2 + Vb2 + 0Vc2 + 0Va3 + Vb3 + 0Vc3 ≤ 1200

0Va1 + 0Vb1 + Vc1 + 0Va2 + 0Vb2 + Vc2 + 0Va3 + 0Vb3 + Vc3 ≤ 750

900Va1 + 900Vb1 + 900Vc1 - 750Va2 - 750Vb2 - 750Vc2 + 0Va3 + 0Vb3 + 0Vc3 = 0

0Va1 + 0Vb1 + 0Vc1 + 450Va2 + 450Vb2 + 450Vc2 - 900Va3 - 900Vb3 - 900Vc3 = 0

450Va1 + 450Vb1 + 450Vc1 + 0Va2 + 0Vb2 + 0Vc2 - 750Va3 - 750Vb3 - 750Vc3 = 0



#for solving it we need to install lpsolve package and call the library of lpsolve

```{r}
library(lpSolve)
```

#Writing the objective function
```{r}
objectivity <- c(420, 360, 300, 420, 360, 300, 420, 360, 300)
objectivity
```


#assigning constraints
```{r}
assign.constraints <- matrix(c(  
  1, 1, 1, 0, 0, 0, 0, 0, 0,  # excess capacity constraints for plant 1
  0, 0, 0, 1, 1, 1, 0, 0, 0,  # excess capacity constraints for plant 2
  0, 0, 0, 0, 0, 0, 1, 1, 1,  # excess capacity constraints for plant 3
  20, 15, 12, 0, 0, 0, 0, 0, 0, # Space constraint for in process storage for plant 1
  0, 0, 0, 20, 15, 12, 0, 0, 0, # Space constraint for in process storage for plant 2
  0, 0, 0, 0, 0, 0, 20, 15, 12, # Space constraint for in process storage for plant 3
  1, 0, 0, 1, 0, 0, 1, 0, 0,  # Forecast of sales constraint for large size
  0, 1, 0, 0, 1, 0, 0, 1, 0,  # Forecast of sales constraint for medium size
  0, 0, 1, 0, 0, 1, 0, 0, 1,   # Forecast of sales constraint for small size
  # use of equal percentage of excess capacity constraint
  900, 900, 900, -750, -750, -750, 0, 0, 0,
  0, 0, 0, 450, 450, 450, -900, -900, -900,
  450, 450, 450, 0, 0, 0, -750, -750, -750), nrow = 12, byrow = TRUE)
assign.constraints
```


# writing the constraints path (<=)
```{r}
path <- c ("<=",
           "<=",
           "<=",
           "<=",
           "<=",
           "<=",
           "<=",
           "<=",
           "<=",
           "=",
           "=",
           "=")
path
```


# Define the right-hand side of constraints
```{r}
rhs <- c(750, 900, 450,        #excess capacity 
         13000, 12000, 5000,   #storage capacity
         900, 1200, 750,       #sales
         0,0,0                 #same % 
         )
rhs
```



```{r}
# Solve the linear programming problem which is maximize profit
lp_result <- lp("max", objectivity, assign.constraints, path, rhs)

# Print the results
lp_result
```


#finding Decision variables values using lp function.
```{r}
lp_result$solution
```

